Supporting Non-Functional Requirements in Services Software Development Process: an MDD Approach

### 2013-10-22 SOFSEM2014/FinalVersion/proofOfConcept.tex
minor changes everywere -- Martin

All the system restrictions are modelled as constraints in this example. 
$\pi$SOD-M provides three types of constraints:
\textit{value}, \textit{business} and \textit{exceptions behaviour} constraints.
Each use case (model) can be associated to one or more constraints\footnote{For a more comprehensive account of $\pi$SOD-M the reader can refer to~\cite{PlacidoThesis2012}.}.
\paragraph*{\textit{$\pi$-UseCase} model:} 
our example has five use cases which represent the
 functions (tasks) and constraints of the system (Figure~\ref{fig:piUC}). 
We do not detail the functional part of the specification, due to lack of space.
The constraints defined are: 
\begin{itemize}
  \item The \textsf{Notify crime} task requires that the user is logged in. 
  This is an example of a \textit{value constraint}, where the value associated to the condition depends on the semantics of the application.
  In this case, it represents the maximum number of allowed login attempts;
\item The \textsf{Store crimes} task requires the verification of the user's clearance (also a value constraint). 
\item In order to perform the \textsf{Track crimes} task,
the contact list of the requesting user must include the user that notified the crime.
This is an example of \textit{business constraint}.
Additionally the requesting user must be logged in.
\item  For the \textsf{View crimes' map} task, the specification defines that if the service \textsf{Google Maps}  is not available, the result is presented as text. This is an example of \textit{exceptional behaviour constraint}. 
The availability of the \textsf{Google Maps} service is verified by a \textit{business constraint}.
 \item The \textsf{Show crimes' details} task is specified to have three constraints: A \textit{value constraint} is defined to verify the user's clearance level; a \textit{business constraint} is used to ensure that the user's clearance is valid for the geographic zone of the crime; another \textit{value constraint} defines that the log is to be maintained.
\end{itemize}
\begin{figure}[t]
\centering
\includegraphics[width=0.8\textwidth]{figs/piUseCase}
\caption{$\pi$-UseCase Model}
\label{fig:piUC}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=1.05\textwidth]{figs/piServiceProcess}
\caption{$\pi$-\textit{ServiceProcess} Model}
\label{fig:piSP}
\end{figure}

@@

\begin{figure}[t]
\centering
\subfloat[$\pi$-ServiceComposition Model]
{\label{fig:piSC}\includegraphics[width=0.95\textwidth]{figs/piServiceComposition}}
~ %add desired spacing between images, e. g. ~, \quad, \qquad etc. (or a blank line to force the subfig onto a new line)
\\[-1mm]
\subfloat[$\pi$-ServiceComposition Policies]
{\label{fig:crimePolicies}\includegraphics[width=0.85\textwidth]{figs/policies}}
~ %add desired spacing between images, e. g. ~, \quad, \qquad etc. (or a blank line to force the subfig onto a new line)
\caption{Service Composition and Policies.}
\label{fig:policies}
\end{figure}
%This model describes the service compositions and their related business
%collaborators (external services), policies, rules associated with a policy and
%the whole application process and functionalities.
%The workflow execution will consider the composition of business services and
%their specific Business Collaborators. 
%Each action will trigger the execution of one or more external services. 
%Before and after each service call, policy rules are verified.

>>>>

All the system restrictions are modelled as constraints in this example. 
$\pi$SOD-M provides three types of constraints:
\textit{value}, \textit{business} and \textit{exceptions behaviour} constraints.
Each use case (model) can be associated to one or more constraints\footnote{For a more comprehensive account of $\pi$SOD-M the reader can refer to~\cite{PlacidoThesis2012}.}.
\paragraph*{\textit{$\pi$-UseCase} model:} 
our example has five use cases which represent the
 functions (tasks) and constraints of the system (Figure~\ref{fig:piUC}). 
We do not detail the functional part of the specification, due to lack of space.
The constraints defined are: 
\begin{itemize}
  \item The \textsf{Notify crime} task requires that the user is logged in. 
  This is an example of a \textit{value constraint}, where the value associated to the condition depends on the semantics of the application.
  In this case, it represents the maximum number of allowed login attempts;
\item The \textsf{Store crimes} task requires the verification of the user's clearance (also a value constraint). 
\item In order to perform the \textsf{Track crimes} task,
the contact list of the requesting user must include the user that notified the crime.
This is an example of \textit{business constraint}.
Additionally the requesting user must be logged in.
\item  For the \textsf{View crimes' map} task, the specification defines that if the service \textsf{Google Maps}  is not available, the result is presented as text. This is an example of \textit{exceptional behaviour constraint}. 
The availability of the \textsf{Google Maps} service is verified by a \textit{business constraint}.
 \item The \textsf{Show crimes' details} task is specified to have three constraints: A \textit{value constraint} is defined to verify the user's clearance level; a \textit{business constraint} is used to ensure that the user's clearance is valid for the geographic zone of the crime; another \textit{value constraint} defines that the log is to be maintained.
\end{itemize}
\begin{figure}[t]
\centering
\includegraphics[width=0.8\textwidth]{figs/piUseCase}
\caption{$\pi$-UseCase Model.}
\label{fig:piUC}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=1.05\textwidth]{figs/piServiceProcess}
\caption{$\pi$-\textit{ServiceProcess} Model.}
\label{fig:piSP}
\end{figure}

@@

\begin{figure}[t]
\centering
\subfloat[$\pi$-ServiceComposition Model.]
{\label{fig:piSC}\includegraphics[width=0.95\textwidth]{figs/piServiceComposition}}
~ %add desired spacing between images, e. g. ~, \quad, \qquad etc. (or a blank line to force the subfig onto a new line)
\\[-1mm]
\subfloat[$\pi$-ServiceComposition Policies.]
{\label{fig:crimePolicies}\includegraphics[width=0.85\textwidth]{figs/policies}}
~ %add desired spacing between images, e. g. ~, \quad, \qquad etc. (or a blank line to force the subfig onto a new line)
\caption{Service Composition and Policies.}
\label{fig:policies}
\end{figure}
%This model describes the service compositions and their related business
%collaborators (external services), policies, rules associated with a policy and
%the whole application process and functionalities.
%The workflow execution will consider the composition of business services and
%their specific Business Collaborators. 
%Each action will trigger the execution of one or more external services. 
%Before and after each service call, policy rules are verified.
